Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-14 thirdPaper/draft/typeConeAssociahedra.tex
schema

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide, float, blkarray, overpic, nicefrac}
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[normalem]{ulem}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}{figures/diagonals/}{figures/walks/}{figures/tubes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\usepackage[export]{adjustbox}

@@

We give a brief overview of our main results and the structure of the article. In particular, some of the geometric realizations in~\cref{part:geometry} depend on algebraic results whose proof is delayed until~\cref{part:algebra}. The goal of this section is to clarify their logical dependencies.

@@

\para{Cluster algebras of finite type}
(\cref{subsec:typeConeCA})

@@

\noindent
The same strategy is applied to a class of finite dimensional algebras, called gentle algebras, whose representations are combinatorially well understood.
We identify two conditions, called brick and 2-acyclicity (\cref{def:brick2acyclic}), that enable us to make use of the type cone approach.
We note that the~$\tau$-tilting theory of gentle algebras is the algebraic counterpart of the combinatorics of accordions on surfaces~\cite{PaluPilaudPlamondon-nonkissing,BrustleDouvilleMousavandThomasYildirim,PaluPilaudPlamondon-surfaces}.

@@

\para{Graphical nested complexes}
(\cref{subsec:typeConeGA})

@@

\noindent
We prove an analogue for cluster categories of a result of M. Auslander for module categories of Artin algebras~\cite{Auslander1984}. 
We are mainly motivated by the fact that it has for a consequence that the type cones of the cluster fans of finite type are simplicial.
However, this result is of independent interest and sheds new lights on the Grothendieck groups of cluster categories.

@@

\noindent
We prove a more general version of \cref{thm:relations-g-vecteurs} that not only applies to cluster fans, but also to non-kissing fans.
Once more, our main motivation is to prove that the type cone of the non-kissing fan of a brick and 2-acyclic gentle algebra is simplicial, but the results are of independent interest, and might be useful for studying the type cones of other families of simplicial fans.

@@

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide, float, blkarray, overpic, nicefrac}
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[normalem]{ulem}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.pathmorphing, decorations.markings, decorations.shapes, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}{figures/diagonals/}{figures/walks/}{figures/tubes/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\usepackage[export]{adjustbox}

@@

We now give a more detailed overview of our main results and the structure of the article. In particular, some of the geometric realizations in~\cref{part:geometry} depend on algebraic results whose proof is delayed until~\cref{part:algebra}. The goal of this section is to clarify their logical dependencies.

@@

\para{Cluster algebras of finite type and generalized associahedra}
(\cref{subsec:typeConeCA})

@@

\noindent
The same strategy is applied to a class of finite dimensional algebras, called gentle algebras, whose representations are combinatorially well understood.
We note that the~$\tau$-tilting theory of gentle algebras is the algebraic counterpart of the combinatorics of accordions on surfaces~\cite{PaluPilaudPlamondon-nonkissing,BrustleDouvilleMousavandThomasYildirim,PaluPilaudPlamondon-surfaces}.
We identify two conditions, called brick and 2-acyclicity (\cref{def:brick2acyclic}), that enable us to make use of the type cone approach.

@@

\para{Graphical nested complexes and graph associahedra}
(\cref{subsec:typeConeGA})

@@

\noindent
We prove an analogue for cluster categories of a result of M. Auslander for module categories of Artin algebras~\cite{Auslander1984}. 
We are mainly motivated by the fact that it implies that the type cones of the cluster fans of finite type are simplicial.
However, this result is of independent interest and sheds new lights on the Grothendieck groups of cluster categories.

@@

\noindent
We prove a more general version of \cref{thm:relations-g-vecteurs} that applies not only to cluster fans, but also to non-kissing fans.
Once more, our main motivation is to prove that the type cone of the non-kissing fan of a brick and 2-acyclic gentle algebra is simplicial, but the results are of independent interest, and might be useful for studying the type cones of other families of simplicial fans.

@@

\noindent
We conclude with a schematic diagram representing the logical dependencies between the different sections of the paper.

@@

\vspace{.5cm}
\tikzset{
  double arrow/.style={
    -latex, line width=3pt, black, % first arrow
    postaction={draw,line width=2pt, white, shorten <= 1,shorten >= 2}, % second arrow
  }
}
\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
	\node[cloud, cloud puffs=25, cloud ignores aspect, minimum width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!25] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!25] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!25] (I1.240) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!25] (I1.300) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!25] (I1.0) to [bend left, looseness=1.1] (I24.90);
	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=(I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}
